\documentclass{review-sheet}
\homework{1}

\begin{document}
\begin{problem}{\S 1}\\
  Translate from set-builder notation into English.
\end{problem}
\begin{solution}
  \begin{enumerate}[label=(\alph*)]
    \item $ \{x\in \R\mid x>0\} \cup \{x\in \R\mid x<0\}$ is the set of all non-zero real numbers.
    \item $\R\setminus \Q$ is the set of all irrational numbers.
    \item $\{\alpha\in \C\mid \alpha^2\in \Q\}$ is the set of all complex numbers whose squares are
      rational.
    \item $\{\left( x, y \right) \in \C^2\mid y=-x\}$ is the set of ordered pairs whose sum is
      zero.
    \item $\{\left( x, 2x \right) \mid x\in \R\}$ is the set of ordered pairs whose second number
      is double the first number.
    \item $\{\left( a, b, 0 \right) \mid a, b\in \R\} \cap \{\left( x, 0, z \right) \mid x, z\in
      \R\} $ is the x axis.
  \end{enumerate}
\end{solution}

\begin{problem}{\S 2}\\
  Translate from English to set-builder notation.
\end{problem}
\begin{solution}
  \begin{enumerate}[label=(\alph*)]
  \item The set of real numbers whose cube is an integer is \[
  \{x\in \R\mid x^3\in \Z\} 
  .\]  
  \item The set of real numbers whose square is greater than $2$ is \[
    \{x\in \R\mid x^2>2\} 
  .\] 
  \item The set of rational numbers strictly between $0$ and $1$ is \[
    \{x\in \Q\mid 0 < x < 1\} 
  .\] 
  \item The set of ordered pairs of complex numbers in which the second is $i$ times the first one
    is \[
      \{\left( \alpha, \beta \right) \in \C^2\mid ai = b\}  
    .\] 
  \item The set of ordered triples of real numbers in which the third one is zero is \[
      \{\left( a, b, c \right) \in \R^3\mid c = 0\} 
    .\] (Alternatively, $ \{\left( a, b, 0 \right)\mid a \in \R \} $).
  \item The set of ordered triples of real numbers in which at least one of the three is zero is \[
      \{\left( a, b, c \right) \in \R^3\mid a = 0 \lor b = 0 \lor c = 0\} 
  .\] 
  \item The set of nonzero complex solutions to the equation $x^{5}+17x^3-x^2=0$ is \[
      \{\alpha \in \C\setminus \left\{ 0 \right\}\mid \alpha^{5} + 17\alpha^{3} - \alpha^2 = 0\}
  .\] 
  \end{enumerate}
\end{solution}

\end{document}
